\texttt{apt-p2p}: A Peer-to-Peer Distribution System for Software Package Releases and Updates

### 2008-08-29 docs/paper/paper.tex
More work on the paper, almost done now.

\begin{abstract}
The Internet has become an cost-effective
vehicle for software development and release, particular in the free software society.
Given the free nature of these software, there are often a number of users
motivated by altruism to help out with the distribution, so as to promote the healthy development
of this voluntary society. It is thus naturally expected that peer-to-peer distribution can be implemented,
which scale well with large user bases and can easily explore the network resources made available by
the volunteers.

Unfortunately, this application scenario has many unique characteristics, which
make a straightforward adoption of such existing peer-to-peer systems for file sharing as BitTorrent suboptimal. In particular,
a software release often consists of a large number of packages, which are difficult to be distributed individually, but the archive is
too large to be distributed in its entirety. The packages are also being constantly
updated by the loosely-managed developed; yet the interest in a particular version of a package can be very
limited depending on the computer platforms and operating systems used.

In this paper, we propose a novel peer-to-peer assisted distributor design that
addresses the above challenges. It enhances the existing distribution systems by providing compatible and yet more efficient downloading and updating services
for software packages. Our design leads to \texttt{apt-p2p}, a practical implementation that extends the popular \texttt{apt} distributor.  \texttt{apt-p2p}  has been used in conjuction with Debian-based distribution of Linux
software packages and will also be available in the next release of Ubuntu. We have addressed the key design issues in \texttt{apt-p2p}, including indexing table customization,
response time reduction, and multi-value extension. They together ensure
that the altruistic users' resources be effectively utilized and thus significantly reduce the currently
sheer bandwidth requirements of hosting the software, as confirmed by our existing real user statistics.
\end{abstract}

@@

With the widespread penetration of broadband accesses, the Internet has become an cost-effective
vehicle for software development and release. This is particularly true for the free software
society whose developers and users are worldwide distributed and work asynchronously. The ever increasing power of
modern programming languages, computer platforms, and operating systems has made these software extremely large and complex,
which are often divided into a huge number of small packages.
Together with their popularity among users,
an efficient and reliable management and distribution of these software over the Internet has become a challenging task.

The existing distributor for free software are mostly based on the client/server model, e.g.,
the Advanced Package Tool (\texttt{apt}) for Linux, which suffers from the well-known bottleneck problem.
Given the free nature of these software, there are often a number of users
motivated by altruism to help out with the distribution, so as to promote the healthy development
of this voluntary society.
We thus naturally expect that peer-to-peer distribution can be implemented in
this context, which scale well with large user bases and can easily explore the network resources made available by
the volunteers.

Unfortunately, this application scenario has many unique characteristics, which
make a straightforward adoption of such existing peer-to-peer systems for file sharing as BitTorrent suboptimal. In particular,
there are too many packages to distribute each individually, but the archive is
too large to distribute in its entirety. The packages are also being constantly
updated by the loosely-managed developed; yet the interest in a particular version of a package can be very
limited. They together make it very difficult to efficiently create and manage torrents/trackers. The random downloading nature of BitTorrent-like systems is also different from the
sequential order used in existing software package distributors. This in turn suppresses interaction with users
given the difficulty in tracking speed and downloading progress.

In this paper, we propose a novel peer-to-peer assisted distributor design that
addresses the above challenges. It enhances the existing distribution systems by providing compatible and yet more efficient downloading and updating services
for software packages. Our design leads to the development of \texttt{apt-p2p}, a practical implementation based on the Debian \footnote{Debian - The Universal Operating System: {\it http://www.debian.org/}} package
distribution system.  We have addressed the key design issues in \texttt{apt-p2p}, including indexing table customization,
response time reduction, and multi-value extension. They together ensure
that the altruistic users' resources be effectively utilized and thus significantly reduce the currently
sheer bandwidth requirements of hosting the software.

\texttt{apt-p2p}  has been used in conjuction with Debian-based distribution of Linux
software packages and will also be available in the next release of Ubuntu. We have evaluated our currently deployment  to
determine how effective it is at meeting our goals, and to see what
effect it is having on the Debian package distribution system. In particular, our existing real user statistics
have suggested that it responsively interacts with clients and substantially reduces server cost.

The rest of this paper is organized as follows. The background and motivation are presented in Section~\ref{situation}, including an analysis of BitTorrent's use for this purpose in Section~\ref{bittorrent}. We propose
our solution in Section~\ref{opportunity}. We then detail our sample
implementation for Debian-based distributions in Section~\ref{implementation},
including an in-depth look at our DHT
customizations in Section~\ref{custom_dht}. The performance of our implementation is evaluated in Section~\ref{analysis}. We examine some related work in Section~\ref{related}, and then
Section~\ref{conclusions} concludes the paper and offers some future directions.

@@

Similar tools have been used for other software packages. CPAN
distributes software packages for the PERL
programming language, using SOAP RPC requests to find and download
files. Cygwin provides many of the
standard Unix/Linux tools in a Windows environment, using a
package management tool that requests packages from websites. There
are two software distribution systems for software that runs on the Macintosh OS, fink and
MacPorts, that also retrieve packages in this way.

@@

Given the free nature of this software, there are often a number of users
motivated by altruism to want to help out with the distribution.
This is particularly true considering that many of this software is used by
groups that are staffed mostly, or sometimes completely, by
volunteers. They are thus motivated to contribute their network resources, so as to promote the healthy development
of the volunteer community that released the software.
We also naturally expect that peer-to-peer distribution can be implementation in
this context, which scale well with large user bases and can easily explore the network resources made available by
the volunteers.

@@

\begin{figure}
\centering
\includegraphics[width=\columnwidth]{apt_p2p_simulation-size_CDF.eps}
\caption{The CDF of the size of packages in a Debian system, both
for the actual size and adjusted size based on the popularity of
the package.}
\label{size_CDF}
\end{figure}

For example, Figure~\ref{size_CDF} shows the size of the packages in the
current Debian distribution. While 80\% of the packages are less than
512~KB, some of the packages are hundreds of megabytes. The entire
archive consists of 22,298 packages and is approximately 119,000 MB
in size. Most of the packages are to be installed in any computer environment, but there are
also OS- or architecture-specific packages.

@@

The software packages being distributed are being constantly
updated. These updates could be the result of the software creators
releasing a new version with improved functionality,
or the distributor updating their packaging of the
software to meet new requirements. Even if the distributor
periodically makes \emph{stable} releases, which are snapshots of
all the packages in the archive at a certain time, updates are still
released for security issues or serious bugs.

\begin{figure}
\centering
\includegraphics[width=\columnwidth]{size-quarter.eps}
\caption{The amount of data in the 119,000 MB Debian archive that is
updated each day, broken down by architecture.}
\label{update_size}
\end{figure}

@@

Finally, though there are a large number of packages and a large number of
users, the interest in a particular version of a package can be very
limited. Specifically, there are core packages that every user has to download, but most
packages would fall in the category of optional or extra, and so are
interesting to only a limited number of people.

\begin{figure}
\centering
\includegraphics[width=\columnwidth]{apt_p2p_popularity-cdf.eps}
\caption{The CDF of the popularity of packages in a Debian system.}
\label{popularity_CDF}
\end{figure}

@@

Given the relatively long time for software package downloading, existing package management systems generally display
some kind of indication of speed and completeness for users to
monitor. Since previous client-server downloads occurred in a sequential
fashion, the package management software also measures the speed
based on sequential downloading. To offer comparable user experience, it is naturally to expect that
the new peer-to-peer solution be reasonably responsive at retrieving packages,
preferably enabling sequential downloading order, too.

@@

An alternative is to create torrents tracking smaller groups of packages. Unfortunately, we find that this enhancement can be
quite difficult given the unique characteristic of free software packages.
First, there is no obvious way to divide the packages into torrents.
Most of the packages are too small, and there are too many packages
in the entire archive to create individual torrents for each one.
On the other hand, all the packages together are too large to track
efficiently as a single torrent. Hence, some division of the archive's
packages into torrents is obviously necessary, but wherever that
split occurs it will cause either some duplication of connections,
or prevent some peers from connecting to others who do have the same
content. In addition, a small number of the packages can be updated every
day which would add new files to the torrent, thereby changing its
\emph{infohash} identifier and making it a new torrent. This will
severely fracture the download population, even though peers in the
new torrent may share 99\% of the packages in common with peers in the
old torrent.

@@

On the other hand, there are aspects of BitTorrent that are no
longer critical. Specifically, with altruistic peers
and all files being available without uploading, incentives to share
become a less important issue. Also, the availability of seeds are
not critical either, as the mirror sites serve in that capacity
already.

@@

We now present the design of our peer-to-peer assisted distributor for free software package releases and
updates. A key principle in our design is that the new functionalities implemented in our distributor should be transparent to users,
thus offering the same experience as using conventional software management systems, despite enhanced efficiency.
That said, we assume that the user is still attempting to download packages from a
server, but the requests will be proxied by our peer-to-peer program. The server is always available and has all of the package files.
In addition, the cryptographic hash of the packages will be available
separately from the package itself, and is usually contained in an
\emph{index} file which also contains all the packages' names,
locations and sizes.

\begin{figure}
\centering
\includegraphics[width=\columnwidth]{model_simple.eps}
\caption{The different phases of functionality of our peer-to-peer distribution model.}
\label{model}
\end{figure}

Our model for using peer-to-peer to enhance package distribution is shown in
Figure~\ref{model}. As shown in Phase~1, our program will act as a
proxy (1,2), downloading (3) and caching all files communicated
between the user and the server (4). It will therefore also have
available the index files containing the cryptographic hashes all
packages. Later, in Phase~2, upon receiving a request from the user
to download a package (5), our program will search the index files
for the package being requested and find its hash (6). This hash can
then be looked up recursively in an indexing structure (a Distributed Hash Table [] in our implementation) (7), which will return a
list of peers that have the package already (8). The package can
then be downloaded from the peers (11,12), it can be verified using
the hash (13), and if valid can be returned to the user (14). The
current nodes location is also added to the DHT for that hash (15),
as it is now a source for others to download from.

In steps (11,12), the fact that this package is also available to download for free
from a server is very important to our proposed model. If the package hash
can not be found in the DHT, the peer can then fallback to
downloading from the original location (i.e. the network of
mirrors). The mirrors thus, with no modification to their
functionality, serve as seeds for the packages in the peer-to-peer
system. Any packages that have just been updated, or that are very
rare, and so don't have any peers available, can always be found on
the mirror. Once the peer has completed the download from the mirror
and verified the package, it can then add itself to the DHT as the
first peer for the new package, so that future requests for the package
will not need the mirror.

This sparse
interest in a large number of packages undergoing constant updating
is well suited to the functionality provided by a Distributed Hash
Table (DHT). DHTs require unique keys to store and retrieve strings
of data, for which the cryptographic hashes used by these package
management systems are perfect for. The stored and retrieved strings
can then be pointers to the peers that have the package that hashes
to that key. A downloading peer can lookup the package hash in the
DHT and, if it is found, download the file from those peers and
verify the package with the hash. Once the download is complete, the
peer will add its entry to the DHT indicating that it now has the
package.

Note that, despite downloading the package from untrustworthy peers,
the trust of the package is always guaranteed through the use
of the cryptographic hashes. Nothing can be downloaded from a peer
until the hash is looked up in the DHT, so a hash must first come
from a trusted source (i.e. the distributors server). Most distributors use index
files that contain hashes for a large number of the packages in
their archive, and which are also hashed. After retrieving the
index's hash from the mirror, the index file can be downloaded from
peers and verified. Then the program has access to all the hashes of
the packages it will be downloading, all of which can be verified
with a \emph{chain of trust} that stretches back to the original
distributor's server.

\subsection{Implementation Options}
\label{imp_options}

There are several ways to implement the desired peer-to-peer functionality
into the existing package management software. The functionality can
be directly integrated through modifications to the software, though
this could be difficult as the peer-to-peer functionality should be running
at all times. This is needed both for efficient lookups and to
support uploading of already downloaded packages. Unfortunately, the
package management tools typically only run until the download and
install request is complete.

Many of the package management software implementations use HTTP
requests from web servers to download the packages, which makes it
possible to implement the peer-to-peer aspect as an almost standard HTTP
caching proxy. This proxy will run as a daemon in the background,
listening for requests from the package management tool for package
files, as well as serving (uploading) cached package to other peers.
It will get uncached requests first from the peer-to-peer system, or
fall back to the normal HTTP request from a server should it not
be found. For methods that don't use HTTP requests, other types of
proxies may also be possible.

\subsection{Downloading From Peers}
\label{downloading}

Although not necessary, we recommend implementing a download
protocol that is similar to the protocol used to fetch packages from
the distributor's servers. This simplifies the peer-to-peer program, as it
can then treat peers and mirrors almost identically when requesting
packages. In fact, the mirrors can be used when there are only a few
slow peers available for a file to help speed up the download
process.

@@

Since the package management system only stores a hash of the entire
package, and not of pieces of that package, we will need to be able
to store and retrieve these piece hashes using the peer-to-peer protocol. In
addition to storing the file download location in the DHT (which
would still be used for small files), a peer will store a
\emph{torrent string} containing the peer's hashes of the pieces of
the larger files. These piece hashes are compared ahead of time
to determine which peers have the same piece hashes (they all
should), and then used during the download to verify the pieces of
the downloaded package.

@@

We created a customized DHT based on Khashmir \cite{khashmir}, which
is an implementation of Kademlia \cite{kademlia}. Khashmir is also the same DHT
implementation used by most of the existing BitTorrent clients to
implement trackerless operation. The communication is all handled by
UDP messages, and RPC (remote procedure call) requests and responses
are all \emph{bencoded} in the same way as BitTorrent's
\texttt{.torrent} files. Khashmir uses the high-level Twisted
event-driven networking engine \cite{twisted}, so we also use
Twisted in our sample implementation for all other networking needs.
More details of this customized DHT can be found below in
Section~\ref{custom_dht}.

@@

\section{Customized DHT / System Optimization}
\label{custom_dht}

A large contribution of our work is in the customization and use of
a Distributed Hash Table (DHT). Although our DHT is based on
Kademlia, we have made many improvements to it to make it suitable
for this application. In addition to a novel storage technique to
support piece hashes, we have improved the response time of looking
up queries, allowed the storage of multiple values for each key, and
incorporated some improvements from BitTorrent's tracker-less DHT
implementation.

\subsection{Kademlia DHT Background}
\label{dht}

DHT's operate by storing (\emph{key}, \emph{value}) pairs in a
distributed fashion such that no node will, on average, store more
than any other node. They support two primitive operations:
\emph{put}, which takes a key and a value and stores it in the DHT;
and \emph{get}, which takes a key and returns a value (or values)
that was previously stored with that key. These operations are
recursive, as each node does not know about all the other nodes in a
DHT, and so must recursively search for the correct node to put to
or get from.

The Kademlia DHT, like most other DHTs, assigns IDs to peers from
the same space that is used for keys. The peers with IDs closest to
the desired key will then store the values for that key. Lookups are
recursive, as nodes are queried in each step that are exponentially
closer to the key than in the previous step.

Nodes in a Kademlia system support four primitive requests.
\texttt{ping} will cause a peer to return nothing, and is only used
to determine if a node is still alive, while \texttt{store} tells a
node to store a value associated with a given key. The most
important primitives are \texttt{find\_node} and
\texttt{find\_value}, which both function recursively to find nodes
close to a key. The queried nodes will return a list of the nodes
they know about that are closest to the key, allowing the querying
node to quickly traverse the DHT to find the nodes close to the
desired key. The only difference between \texttt{find\_node} and \texttt{find\_value} is that the
\texttt{find\_value} query will cause a node to return a value, if
it has one for that key, instead of a list of nodes.

@@

\begin{figure}
\centering
\includegraphics[width=\columnwidth]{apt_p2p_improvements-find_value.eps}
\caption{The distribution of average response times PlanetLab nodes
experience for \texttt{find\_value} queries. The original DHT
implementation results are shown, as well as the successive
improvements that we made to reduce the response time.}
\label{improvements}
\end{figure}

To test our changes during development, we ran the customized DHT
for several hours after each major change on over 300 PlanetLab nodes
\cite{planetlab}. Though the nodes are not expected to be firewalled
or NATted, some can be quite overloaded and so consistently fail to
respond within a timeout period, similar to NATted peers. The
resulting distribution of the nodes' average response times is shown
in Figure~\ref{improvements}. Each improvement successfully reduced
the response time, for a total reduction of more than half. The
final distribution is also narrower, as the improvements make the
system more predictable. However, there are still a large number of
outliers with higher average response times, which are the
overloaded nodes on PlanetLab. This was confirmed by examining the
average time it took for a timeout to occur, which should be
constant as it is a configuration option, but can be much larger if
the node is too overloaded for the program to be able to check for a
timeout very often.

\subsection{Extension with Multiple Values}
\label{multiple_values}

@@

@@

\begin{figure}
\centering
\includegraphics[width=\columnwidth]{AptP2PWalker-peers.eps}
\caption{The number of peers found in the system, and how many are
behind a firewall or NAT.}
\label{walker_peers}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics[width=\columnwidth]{AptP2PDuration-peers.eps}
\caption{The CDF of how long an average session will last.}
\label{duration_peers}
\end{figure}

>>>>

\begin{abstract}
The Internet has become a cost-effective
vehicle for software development and release, particular in the free software community.
Given the free nature of these software, there are often a number of users
motivated by altruism to help out with the distribution, so as to promote the healthy development
of this voluntary society. It is thus naturally expected that peer-to-peer distribution can be implemented,
which scale well with large user bases and can easily explore the network resources made available by
the volunteers.

Unfortunately, this application scenario has many unique characteristics, which
make a straightforward adoption of existing peer-to-peer systems for file sharing such as BitTorrent suboptimal. In particular,
a software release often consists of a large number of packages, which are difficult to be distributed individually, but the archive is
too large to be distributed in its entirety. The packages are also being constantly
updated by the loosely-managed developers, and the interest in a particular version of a package can be very
limited depending on the computer platforms and operating systems used.

In this paper, we propose a novel peer-to-peer assisted distribution system design that
addresses the above challenges. It enhances the existing distribution systems by providing compatible and yet more efficient downloading and updating services
for software packages. Our design leads to \texttt{apt-p2p}, a practical implementation that extends the popular \texttt{apt} distributor.  \texttt{apt-p2p}  has been used in conjuction with Debian-based distribution of Linux
software packages and will also be available in the next release of Ubuntu. We have addressed the key design issues in \texttt{apt-p2p}, including indexing table customization,
response time reduction, and multi-value extension. They together ensure
that the altruistic users' resources are effectively utilized and thus significantly reduces the currently
large bandwidth requirements of hosting the software, as confirmed by our existing real user statistics.
\end{abstract}

@@

With the widespread penetration of broadband access, the Internet has become a cost-effective
vehicle for software development and release. This is particularly true for the free software
community whose developers and users are distributed worldwide and work asynchronously. The ever increasing power of
modern programming languages, computer platforms, and operating systems has made this software extremely large and complex,
though it is often divided into a huge number of small packages.
Together with their popularity among users,
an efficient and reliable management and distribution of these software packages over the Internet has become a challenging task.

The existing distribution for free software is mostly based on the client/server model, e.g.,
the Advanced Package Tool (\texttt{apt}) for Linux, which suffers from the well-known bottleneck problem.
Given the free nature of this software, there are often a number of users
motivated by altruism to help out with the distribution, so as to promote the healthy development
of this voluntary society.
We thus naturally expect that peer-to-peer distribution can be implemented in
this context, which will scale well with the currently large user bases and can easily explore the network resources made available by
the volunteers.

Unfortunately, this application scenario has many unique characteristics, which
make a straightforward adoption of existing peer-to-peer systems for file sharing such as BitTorrent suboptimal. In particular,
there are too many packages to distribute each individually, but the archive is
too large to distribute in its entirety. The packages are also being constantly
updated by the loosely-managed developers, and the interest in a particular version of a package can be very
limited. They together make it very difficult to efficiently create and manage torrents/trackers. The random downloading nature of BitTorrent-like systems is also different from the
sequential order used in existing software package distributions. This in turn suppresses interaction with users
given the difficulty in tracking speed and downloading progress.

In this paper, we propose a novel peer-to-peer assisted distribution system design that
addresses the above challenges. It enhances the existing distribution systems by providing compatible and yet more efficient downloading and updating services
for software packages. Our design leads to the development of \texttt{apt-p2p}, a practical implementation based on the Debian \footnote{Debian - The Universal Operating System: {\it http://www.debian.org/}} package
distribution system.  We have addressed the key design issues in \texttt{apt-p2p}, including indexing table customization,
response time reduction, and multi-value extension. They together ensure
that the altruistic users' resources are effectively utilized and thus significantly reduces the currently
large bandwidth requirements of hosting the software.

\texttt{apt-p2p}  has been used in conjuction with Debian-based distribution of Linux
software packages and will also be available in the next release of Ubuntu. We have evaluated our current deployment  to
determine how effective it is at meeting our goals, and to see what
effect it is having on the Debian package distribution system. In particular, our existing real user statistics
have suggested that it responsively interacts with clients and substantially reduces server cost.

The rest of this paper is organized as follows. The background and motivation are presented in Section~\ref{situation}, including an analysis of BitTorrent's use for this purpose in Section~\ref{bittorrent}. We propose
our solution in Section~\ref{opportunity}. We then detail our sample
implementation for Debian-based distributions in Section~\ref{implementation},
including an in-depth look at our system optimization
in Section~\ref{custom_dht}. The performance of our implementation is evaluated
in Section~\ref{analysis}. We examine some related work in Section~\ref{related}, and then
Section~\ref{conclusions} concludes the paper and offers some future directions.

@@

Similar tools have been used for other types of software packages. CPAN
distributes packaged software for the PERL
programming language, using SOAP RPC requests to find and download
files. Cygwin provides many of the
standard Unix/Linux tools in a Windows environment, using a
package management tool that requests packages from websites. There
are two software distribution systems for software that runs on the Macintosh OS, fink and
MacPorts, that also retrieve packages in this way.

@@

Given the free nature of this software, there are often a number of users
motivated by altruism to want to help out with the distribution.
This is particularly true considering that much of this software is used by
groups that are staffed mostly, or sometimes completely, by
volunteers. They are thus motivated to contribute their network resources, so as to promote the healthy development
of the volunteer community that released the software.
We also naturally expect that peer-to-peer distribution can be implemented in
this context, which will scale well with the currently large user bases
and can easily explore the network resources made available by
the volunteers.

@@

\begin{figure}
\centering
\includegraphics[width=0.80\columnwidth]{apt_p2p_simulation-size_CDF.eps}
\caption{The CDF of the size of packages in a Debian system, both
for the actual size and adjusted size based on the popularity of
the package.}
\label{size_CDF}
\end{figure}

For example, Figure~\ref{size_CDF} shows the size of the packages in the
current Debian distribution. While 80\% of the packages are less than
512~KB, some of the packages are hundreds of megabytes. The entire
archive consists of 22,298 packages and is approximately 119,000 MB
in size. Many of the packages are to be installed in any computer environment, but there are
also OS- or architecture-specific packages.

@@

The software packages being distributed are being constantly
updated. These updates could be the result of the software creators
releasing a new version with improved functionality,
or the distributor updating their packaging of the
software to meet new requirements. Even if the distributor
periodically makes \emph{stable} releases, which are snapshots of
all the packages in the archive at a certain time, later updates are still
released for security issues or serious bugs.

\begin{figure}
\centering
\includegraphics[width=0.9\columnwidth]{size-quarter.eps}
\caption{The amount of data in the 119,000 MB Debian archive that is
updated each day, broken down by architecture.}
\label{update_size}
\end{figure}

@@

Finally, though there are a large number of packages and a large number of
users, the interest in a particular package, or version of a package, can be very
limited. Specifically, there are core packages that every user has to download, but most
packages would fall in the category of optional or extra, and so are
interesting to only a limited number of people.

\begin{figure}
\centering
\includegraphics[width=0.80\columnwidth]{apt_p2p_popularity-cdf.eps}
\caption{The CDF of the popularity of packages in a Debian system.}
\label{popularity_CDF}
\end{figure}

@@

Given the relatively long time for software package downloading, existing package management systems generally display
some kind of indication of speed and completeness for users to
monitor. Since previous client-server downloads occurred in a sequential
fashion, the package management software also measures the speed
based on sequential downloading. To offer comparable user experience, it is natural to expect that
the new peer-to-peer solution be reasonably responsive at retrieving packages,
preferably in a sequential downloading order too.

@@

An alternative is to create torrents tracking smaller groups of packages. Unfortunately, we find that this enhancement can be
quite difficult given the unique characteristic of free software packages.
First, there is no obvious way to divide the packages into torrents.
Most of the packages are too small, and there are too many packages
in the entire archive, to create individual torrents for each one.
On the other hand, all the packages together are too large to track
efficiently as a single torrent. Hence, some division of the archive's
packages into torrents is obviously necessary, but wherever that
split occurs it will cause either some duplication of connections,
or prevent some peers from connecting to others who do have the same
content. In addition, a small number of the packages can be updated every
day which would add new files to the torrent, thereby changing its
\emph{infohash} identifier and making it a new torrent. This will
severely fracture the download population, since even though peers in the
new torrent may share 99\% of the packages in common with peers in the
old torrent, they will be unable to communicate.

@@

On the other hand, there are aspects of BitTorrent that are no
longer critical. Specifically, with altruistic peers
and all files being available to download without uploading, incentives to share
become a less important issue. Also, the availability of seeds are
not critical either, as the serves are already available to serve in
that capacity.

@@

We now present the design of our peer-to-peer assisted distribution system for free software package releases and
updates. A key principle in our design is that the new functionalities implemented in our distributor should be transparent to users,
thus offering the same experience as using conventional software management systems, but with enhanced efficiency.
That said, we assume that the user is still attempting to download packages from a
server, but the requests will be proxied by our peer-to-peer program.
We further assume that the server is always available and has all of the package files.
In addition, the cryptographic hash of the packages will be available
separately from the package itself, and is usually contained in an
\emph{index} file which also contains all the packages' names,
locations and sizes.

\begin{figure}
\centering
\includegraphics[width=0.9\columnwidth]{model_simple.eps}
\caption{The different phases of functionality of our peer-to-peer distribution model.}
\label{model}
\end{figure}

Our model for using peer-to-peer to enhance package distribution is shown in
Figure~\ref{model}. As shown in Phase~1, our program will act as a
proxy (1,2), downloading (3) and caching all files communicated
between the user and the server (4). It will therefore also have
available the index files containing the cryptographic hashes all
packages. Later, in Phase~2, upon receiving a request from the user
to download a package (5), our program will search the index files
for the package being requested and find its hash (6). This hash can
then be looked up recursively in an indexing structure (a Distributed Hash Table in our implementation) (7), which will return a
list of peers that have the package already (8). The package can
then be downloaded from the peers (11,12), it can be verified using
the hash (13), and if valid can be returned to the user (14). The
current nodes location is also added to the DHT for that hash (15),
as it is now a source for others to download from.

In steps (11,12), the fact that this package is also available to download for free
from a server is very important to our proposed model. If the package hash
can not be found in the DHT, the peer can then fallback to
downloading from the original location (i.e. the server).
The server thus, with no modification to its
functionality, serves as a seed for the packages in the peer-to-peer
system. Any packages that have just been updated, or that are very
rare, and so don't have any peers available, can always be found on
the server. Once the peer has completed the download from the server
and verified the package, it can then add itself to the DHT as the
first peer for the new package, so that future requests for the package
will not need the server.

This sparse
interest in a large number of packages undergoing constant updating
is well suited to the functionality provided by a Distributed Hash
Table (DHT). DHTs require unique keys to store and retrieve strings
of data, for which the cryptographic hashes used by these package
management systems are perfect for. The stored and retrieved strings
can then be pointers to the peers that have the package that hashes
to that key.

Note that, despite downloading the package from untrustworthy peers,
the trust of the package is always guaranteed through the use
of the cryptographic hashes. Nothing can be downloaded from a peer
until the hash is looked up in the DHT, so a hash must first come
from a trusted source (i.e. the distributors server). Most distributors use index
files that contain hashes for a large number of the packages in
their archive, and which are also hashed. After retrieving the
index's hash from the server, the index file can be downloaded from
peers and verified. Then the program has access to all the hashes of
the packages it will be downloading, all of which can be verified
with a \emph{chain of trust} that stretches back to the original
distributor's server.

% \subsection{Implementation Options}
% \label{imp_options}
% 
% There are several ways to implement the desired peer-to-peer functionality
% into the existing package management software. The functionality can
% be directly integrated through modifications to the software, though
% this could be difficult as the peer-to-peer functionality should be running
% at all times. This is needed both for efficient lookups and to
% support uploading of already downloaded packages. Unfortunately, the
% package management tools typically only run until the download and
% install request is complete.
% 
% Many of the package management software implementations use HTTP
% requests from web servers to download the packages, which makes it
% possible to implement the peer-to-peer aspect as an almost standard HTTP
% caching proxy. This proxy will run as a daemon in the background,
% listening for requests from the package management tool for package
% files, as well as serving (uploading) cached package to other peers.
% It will get uncached requests first from the peer-to-peer system, or
% fall back to the normal HTTP request from a server should it not
% be found. For methods that don't use HTTP requests, other types of
% proxies may also be possible.

\subsection{Peer Downloading Protocol}
\label{downloading}

Although not necessary, we recommend implementing a download
protocol that is similar to the protocol used to fetch packages from
the distributor's server. This simplifies the peer-to-peer program, as it
can then treat peers and the server almost identically when requesting
packages. In fact, the server can be used when there are only a few
slow peers available for a file to help speed up the download
process.

@@

Since the package management system only stores a hash of the entire
package, and not of pieces of that package, we will need to be able
to store and retrieve these piece hashes using the peer-to-peer protocol. In
addition to storing the file download location in the DHT (which
would still be used for small files), a peer will store a
\emph{torrent string} containing the peer's hashes of the pieces of
the larger files (similar to (15) in Phase 3 of Figure~\ref{model}).
These piece hashes will be retrieved and compared ahead of time by
the downloading peer ((9,10) in Phase 2 of Figure~\ref{model})
to determine which peers have the same piece hashes (they all
should), and then used during the download to verify the pieces of
the downloaded package.

@@

We created a customized DHT based on Khashmir \cite{khashmir}, which
is an implementation of Kademlia \cite{kademlia}. Khashmir is also the same DHT
implementation used by most of the existing BitTorrent clients to
implement trackerless operation. The communication is all handled by
UDP messages, and RPC (remote procedure call) requests and responses
are all \emph{bencoded} in the same way as BitTorrent's
\texttt{.torrent} files.
% Khashmir uses the high-level Twisted
% event-driven networking engine \cite{twisted}, so we also use
% Twisted in our sample implementation for all other networking needs.
More details of this customized DHT can be found below in
Section~\ref{custom_dht}.

@@

\section{System Optimization}
\label{custom_dht}

Another contribution of our work is in the customization and use of
a Distributed Hash Table (DHT). Although our DHT is based on
Kademlia, we have made many improvements to it to make it suitable
for this application. In addition to a novel storage technique to
support piece hashes, we have improved the response time of looking
up queries, allowed the storage of multiple values for each key, and
incorporated some improvements from BitTorrent's tracker-less DHT
implementation.

\subsection{DHT Details}
\label{dht}

DHTs operate by storing (\emph{key}, \emph{value}) pairs in a
distributed fashion such that no node will, on average, store more or have to work harder
than any other node. They support two primitive operations:
\emph{put}, which takes a key and a value and stores it in the DHT;
and \emph{get}, which takes a key and returns a value (or values)
that was previously stored with that key. These operations are
recursive, as each node does not know about all the other nodes in the
DHT, and so must recursively search for the correct node to put to
or get from.

The Kademlia DHT, like most other DHTs, assigns IDs to peers randomly from
the same space that is used for keys. The peers with IDs closest to
the desired key will then store the values for that key.
Nodes support four primitive requests.
\texttt{ping} will cause a peer to return nothing, and is only used
to determine if a node is still alive. \texttt{store} tells a
node to store a value associated with a given key. The most
important primitives are \texttt{find\_node} and
\texttt{find\_value}, which both function recursively to find nodes
close to a key. The queried nodes will return a list of the nodes
they know about that are closest to the key, allowing the querying
node to quickly traverse the DHT to find the nodes closest to the
desired key. The only difference between \texttt{find\_node} and \texttt{find\_value} is that the
\texttt{find\_value} query will cause a node to return a value, if
it has one for that key, instead of a list of nodes.

@@

\begin{figure}
\centering
\includegraphics[width=0.80\columnwidth]{apt_p2p_improvements-find_value.eps}
\caption{The distribution of average response times PlanetLab nodes
experience for \texttt{find\_value} queries. The original DHT
implementation results are shown, as well as the successive
improvements that we made to reduce the response time.}
\label{improvements}
\end{figure}

To test our changes during development, we ran our customized DHT
for several hours after each major change on over 300 PlanetLab nodes
\cite{planetlab}. Though the nodes are not expected to be firewalled
or NATted, some can be quite overloaded and so consistently fail to
respond within a timeout period, similar to NATted peers. The
resulting distribution of the nodes' average response times is shown
in Figure~\ref{improvements}. Each improvement successfully reduced
the response time, for a total reduction of more than 50\%. The
final distribution is also narrower, as the improvements make the
system more predictable. However, there are still a large number of
outliers with higher average response times, which are the
overloaded nodes on PlanetLab. This was confirmed by examining the
average time it took for a timeout to occur, which should be
constant as it is a configuration option, but can be much larger if
the node is too overloaded for the program to be able to check for a
timeout very often.

\subsection{Multiple Values Extension}
\label{multiple_values}

@@

% \subsection{\textbf{OPTIONAL}: BitTorrent's Improvements}
% \label{bittorrent_dht}
% 
% In the several years that some BitTorrent clients have been using a
% Kademlia-based DHT for tracker-less operation, the developers have
% made many enhancements which we can take advantage of. One of the
% most important is a security feature added to stop malicious nodes
% from subscribing other nodes as downloaders. When a node issues a
% request to another node to store its download info for that key, it
% must include a \emph{token} returned by the storing node in a recent
% \texttt{find\_node} request. The token is created by hashing the IP
% address of the requesting peer with a temporary secret that expires
% after several minutes. This prevents the requesting peer from faking
% its IP address in the store request, since it must first receive a
% response from a \texttt{find\_node} on that IP.
% 
% We also made some BitTorrent-inspired changes to the parameters of
% the DHT originally specified by the authors of Kademlia. Since, as
% we will show later, peers stay online in our system for much longer
% periods of time, we reduced Kademlia's \emph{k} value from 20 to 8.
% The value is supposed to be large enough such that any given
% \emph{k} nodes are unlikely to fail within an hour of each other,
% which is very unlikely in our system given the long uptimes of
% nodes (see Figure~\ref{duration_online_1} in Section~\ref{analysis}.
% We also increased the number of concurrent outstanding
% requests allowed from 3 to 6 to speed up the recursive key finding
% processes.
% 
% \subsection{\textbf{OPTIONAL}: Other Changes}
% \label{other_changes}
% 
% In addition, we have allowed peers to store values in the DHT, even
% if the hash they are using is not the correct length. Most of the
% keys used in the DHT are based on the SHA1 hash, and so they are 20
% bytes in length. However, some files in the targetted Debian package
% system are only hashed using the MD5 algorithm, and so the hashes
% retrieved from the server are 16 bytes in length. The DHT will still
% store values using these keys by padding the end of them with zeroes
% to the proper length, as the loss of uniqueness from the 4 less
% bytes is not an issue. Also, the DHT will happily store hashes that
% are longer than 20 bytes, such as those from the 32 byte SHA256
% algorithm, by truncating them to the correct length. Since the
% requesting peer will have the full length of the hash, this will not
% affect its attempt to verify the downloaded file.

@@

\begin{figure}
\centering
\includegraphics[width=0.80\columnwidth]{AptP2PWalker-peers.eps}
\caption{The number of peers found in the system, and how many are
behind a firewall or NAT.}
\label{walker_peers}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics[width=0.80\columnwidth]{AptP2PDuration-peers.eps}
\caption{The CDF of how long an average session will last.}
\label{duration_peers}
\end{figure}
